---
title: "Stochastic Processes"
author: "Sami Cheong"
date: "11/4/2020"
output: html_document
---

## Example of a spatial Poisson process 

```{r,message=FALSE,warning=FALSE}


sp_sim <-function(lambda,a_x, a_y){
  ## INPUT:
  # lambda : intensity of arrival
  # a_x = width of the rectangle (x-dim)
  # a_y = length of the rectange (y-dim)
  ## OUTPUT:
  # process$x <- x coordinates of the spatial process
  # process$y <- y coordinates of the spatial process
  # process$n <- number of points generated for the spatial process
  
  # define area of rectangle
  A <- a_x*a_y
  # initiate output with empty list 
  process <-list()
  
  # simulate number of events with rate lambda*area
  N <-rpois(1,lambda*A)
  # simulate uniformly distributed points (x-dim)
  x <-runif(N,min = 0,max = a_x)
  # simulate uniformly distributed points (y-dim)
  y <-runif(N,min = 0,max = a_y)
  
  process$x <-x
  process$y <-y
  process$n <-N
  
  
  return(process)
}



P <-sp_sim(lambda = 10,3,5)
plot(P, main = 'A realization of a Spatial Poisson process \n with rate 10 an on 3-by-5 region', ylab ='y', xlab = 'x')

```


Alternatively, we can also use the `spatstat` package to simulate a spatial Poisson process, below is an example on a unit square:

```{r,message=FALSE}


require(spatstat)



```

### Determining if a spatial point process is completely random


```{r}
# simulate a spatial poisson process with rate lambda = 100
X <- spatstat::rpoispp(100)

# plot X
plot(X, main = 'spatial poisson process with lambda = 100')

# plot Ripley's K function for X with edge correction to fix bias
K_X <-Kest(X,correction = 'best')
plot(K_X)

```

### Deteremine the expected number of points of multiple type 

```{r,message=FALSE}

Kcross_hamster<- Kcross(hamster, "dividing", "pyknotic",correction = 'best') 
plot(Kcross_hamster)

```



### Example of a Non-homogenous Poisson process

Consider the intensity function that represent the popular times of a restaurant

```{r,message=FALSE}
lambda_t<-function(time_int){
  l <-numeric(length(time_int))
  i<-1
  for(t in time_int){
    
    if(t>=0&t<=2){
      
      l[i] <- 10+30*t
      
    }
    
    if(t>2 & t<4){
      
      l[i] <-70
    }
    
    if(t>=4&t<=6){
      l[i]<-70-30*(t-4)
      
    }
    
    if(t>6&t<=8){
      l[i]<-10+30*(t-6)
    }
   
    if(t>8 & t <=10){
      l[i]<-70
    }
    
    if(t>10&t<=12){
      l[i]<-70-30*(t-10)
    }
    i<-i+1
  }
   return(l)
}
time_int <- seq(0,12,0.1)

plot(time_int,lambda_t(time_int),type = 'l',ylab = 'customers per hour', xlab='huors since opening',main = 'Example of lambda(t)' )

# integrate intensity function betweeen 3- 5
lambda_rate <-integrate(lambda_t,3,5)
# probability P(N_t > 100)
k<-100
1-sum(dpois(0:(k-1),lambda_rate$value))


```


